### 2014-12-16 revision/Paper_v3.tex
Final revesion from JV

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large Dynamic Modeling of Cell-Free Biochemical Networks using Effective Kinetic Models}}
\vspace{0.05in}
{\par \centering \large{Joseph A. Wayman, Adithya Sagar and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering \large{School of Chemical and Biomolecular Engineering}}
{\par \centering \large{Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~Effective models of metabolism}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\textit{Processes}}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Associate Professor, School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853} 
{\par \centering Email: jdv27@cornell.edu} 
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Cell-free systems offer many advantages for the study, manipulation and modeling of metabolism compared to \textit{in vivo} processes.
Many of the challenges confronting genome-scale kinetic modeling can potentially be overcome in a cell-free system. 
For example, there is no complex transcriptional regulation to consider, transient metabolic measurements are easier to obtain, and we no longer have to consider cell growth. 
Thus, cell-free operation holds several significant advantages for model development, identification and validation. 
Theoretically, genome-scale cell-free kinetic models may be possible for industrially important organisms, such as \textit{E. coli}, if a simple, tractable framework for integrating allosteric regulation with enzyme kinetics can be formulated. 
Toward this unmet need, we present an effective biochemical network modeling framework for building dynamic cell-free metabolic models.
The key innovation of our approach is the integration of simple effective rules encoding complex allosteric regulation with traditional kinetic pathway modeling. 
We tested our approach by modeling the time evolution of several hypothetical cell-free metabolic networks. 
We found that simple effective rules, when integrated with traditional enzyme kinetic expressions, captured complex allosteric patterns such as ultrasensitivity or non-competitive inhibition in the absence of mechanistic information. 
Second, when integrated into network models, these rules captured classic regulatory patterns such as product-induced feedback inhibition. 
Lastly, we showed, at least for the network architectures considered here, that we could simultaneously estimate kinetic parameters and allosteric connectivity from synthetic data. 
While only an initial proof-of-concept, the framework presented here could be an important first step toward genome-scale cell-free kinetic modeling of the biosynthetic capacity of industrially important organisms.

@@

\section*{Introduction}
Mathematical modeling has long contributed to our understanding of metabolism. 
Decades before the genomics revolution, mechanistically, structured metabolic models arose from the desire to predict microbial phenotypes resulting from changes in intracellular or extracellular states \citep{1976_fredrickson_BiotechBioeng}. 
The single cell \textit{E. coli} models of Shuler and coworkers pioneered the construction of large-scale, dynamic metabolic models that incorporated multiple, regulated catabolic and anabolic pathways constrained by experimentally determined kinetic parameters \citep{1984_domach_shuler_BiotechBioeng_01}. 
Shuler and coworkers generated many single cell kinetic models, including single cell models of eukaryotes \citep{1989_steinmeyer_shuler_ChemEngSci,1992_wu_shuler_AnnNYAcadSci}, minimal cell architectures \citep{2004_castellanos_shuler_PNAS}, as well as DNA sequence based whole-cell models of \textit{E. coli} \citep{2008_atlas_shuler_IETSysBio}.
Conversely, highly abstracted kinetic frameworks, such as the cybernetic framework, represented a paradigm shift, viewing cells as growth-optimizing strategists \citep{1985_dhurjati_ramkrishna_tsao_BiotechBioeng}. 
Cybernetic models have been highly successful at predicting metabolic choice behavior, e.g., diauxie behavior \citep{1986_kompala_ramkrishna_tsao_BiotechBioeng}, steady-state multiplicity \citep{2012_kim_ramkrishna_BiotechProg}, as well as the cellular response to metabolic engineering modifications \citep{1999_varner_ramkrishna_MetaEng}. 
Unfortunately, traditional, fully structured cybernetic models also suffer from an identifiability challenge, as both the kinetic parameters and an abstracted model of cellular objectives must be estimated simultaneously. 
Recent cybernetic formulations from Ramkrishna and colleagues have successfully treated this identifiability challenge through elementary mode reduction, though the techniques replace detailed biological mechanism with an optimization heuristic \cite{2009_song_ramkrishna_BiotechBioeng,Song:2011aa}.

@@

In this study, we present an effective biochemical network modeling framework for building dynamic cell-free metabolic models.
The key innovation of our approach is the seamless integration of simple effective rules encoding complex regulation with traditional kinetic pathway modeling. 
This integration allows the description of complex regulatory interactions, such as time-dependent allosteric regulation of enzyme activity, in the absence of specific mechanistic information. 
The regulatory rules are easy to understand, easy to formulate and do not rely on overarching theoretical abstractions or restrictive assumptions. 
We tested our approach by modeling the time evolution of several hypothetical cell-free metabolic networks. 
In particular, we tested whether our effective modeling approach could describe classically expected enzyme kinetic behavior, and second whether we could simultaneously estimate kinetic parameters and regulatory connectivity, in the absence of specific mechanistic knowledge, from synthetic experimental data. 
Toward these questions, we explored five hypothetical cell-free networks. 
Each network shared the same enzymatic connectivity, but had different allosteric regulatory connectivity. 
We found that simple effective rules, when integrated with traditional enzyme kinetic expressions, captured complex allosteric patterns such as ultrasensitivity or non-competitive inhibition in the absence of mechanistic information. 
Second, when integrated into network models, these rules captured classical regulatory patterns such as product-induced feedback inhibition. 
Lastly, we showed, at least for the network architectures considered here, that we could simultaneously estimate kinetic parameters and allosteric connectivity from synthetic data. 
While only an initial proof-of-concept, the framework presented here could be an important first step toward genome-scale cell-free kinetic modeling of the biosynthetic capacity of industrially important organisms.

@@

We modified our particle swarm identification strategy to simultaneously search over both kinetic parameters and putative control structures. 
In addition to our initial networks, we constructed three additional presumptive network models, each with the same enzymatic connectivity but different allosteric regulation of the pathway enzymes (Fig. \ref{fig-alternative-networks}). 
We then initialized a population of particles, each with one of the five potential regulatory programs and randomized kinetic parameters. 
Thus, we generated an initial population of particles that had \textit{both} different kinetic parameters as well as different control structures. 
We biased the distribution of the particle population according to our \textit{a prior} belief of the correct regulatory program. 
To this end, we considered three different priors, a uniform distribution where each putative regulatory structure represented 20\% of the population and two mixed distributions that were either positively or negatively biased towards the correct structure (network A). 
In both the positively biased and uniform cases the PSO clearly differentiated between the true or closely related structures and those that were materially different (Fig. \ref{fig-control-search}). 
As expected, the positively biased population (40\% of the initial particle population seeded with network A) gave the best results, where the correct structure was preferentially identified (Fig. \ref{fig-control-search}A). 
On the other hand, when given a uniform distribution, the PSO approach identified a combination of network A and network C as the most likely control structures (Fig. \ref{fig-control-search}B). 
Network A and C differ by the regulatory connection between the end product P$_2$ and enzyme $E_{1}$; in network A, end product P$_{2}$ was assumed to inhibit $E_{1}$, while in network C, end product P$_{2}$ activated $E_{1}$.  
Lastly, when the initial population was biased towards incorrect structures (initial population seeded with 90\% incorrect structures), the particle swarm \textit{misidentified} the correct allosteric structure (Fig. \ref{fig-control-search}C). 
Interestingly, while each particle swarm identified parameter sets that minimized the simulation error, the estimated parameter values were not necessarily similar to the true parameters. 
The angle between the estimated and true parameters was not consistently small across the swarms (identical parameters would give an angle of zero). 
This suggested that our particle swarm approach identified a \textit{sloppy} ensemble, i.e., parameter estimates that were individually incorrect but collectively exhibited the correct model behavior.

@@

\section*{Discussion}
In this study, we presented an effective kinetic modeling strategy to dynamically simulate cell-free biochemical networks. 
Our proposed strategy integrated traditional kinetic modeling with an effective rules based approach to dynamically describe metabolic regulation and control. 
We tested this approach by developing kinetic models of hypothetical cell-free metabolic networks. 
In particular, we tested whether our effective modeling approach could describe classically expected behavior, and second whether we could simultaneously estimate kinetic parameters and regulatory connectivity, in the absence of specific mechanistic knowledge, from synthetic experimental data. 
Toward these questions, we explored five hypothetical cell-free networks. 
In each network, a substrate S was converted to the end products P$_{1}$ and P$_{2}$ through a series of enzymatically catalyzed reactions, including a branch point at a hypothetical metabolite M$_{2}$. 
Each network also included the same cofactors and cofactor recycle architecture. 
However, while all five networks shared the same enzymatic connectivity, each had different allosteric regulatory connectivity. 
We found that simple effective rules, when integrated with traditional enzyme kinetic expressions, could capture complex allosteric patterns such as ultrasensitivity, or non-competitive inhibition in the absence of specific mechanistic information. 
Moreover, when integrated into network models, these rules captured classical regulatory patterns such as product-induced feedback inhibition. 
Lastly, we simultaneously estimated kinetic parameters and discriminated between competing regulatory structures, using synthetic data in combination with a modified particle swarm approach.
If we considered all putative regulatory architectures to be equally likely, we were able to estimate a \textit{sloppy} ensemble of models with the correct architecture and kinetic parameters.

@@

There are several critical questions that should be explored following this proof-of-concept study.
It is unclear how parameter identification will scale to genome-scale networks, and second it is unclear how we will identify allosteric connectivity at a genome-scale. 
The enzymatic connectivity for genome-scale cell-free networks can easily be established by stripping away the growth and cell wall machinery from whole cell genome reconstructions.
Then metabolic fluxes can be transformed into kinetic expressions using heuristics such multiple saturation kinetics, which are then modified by our rule-based control variables.  
This leaves a large number of unknown kinetic constants that must be estimated from time-resolved metabolite measurements. 
Ensemble modeling is a well-established approach for parameter identification in large-scale deterministic models. 
Liao and coworkers developed a method that generates an ensemble of kinetic models that all approach the same steady-state, one determined by fluxomics measurements \cite{Tran:2008aa}. 
The best subpopulation of candidate models are selected based on their agreement with further measurements of genetically perturbed systems. 
Our work relies on heuristic search optimization to identify kinetic models consistent with steady-state and dynamic time-series measurements of cellular species \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa,Lequieu:2011aa}. 
Instead of estimating a single yet highly uncertain parameter set, both approaches estimate an ensemble of parameter sets whose model behavior recapitulates experimental measurements. 
Here, we showed that particle swarm optimization quickly identified an ensemble of model parameters, at least for proof-of-concept metabolic networks using synthetic data. 
This suggested that we can expect reasonable model predictions, despite only partial parameter knowledge, as network size grows if we have properly designed experiments. 
Brown and Sethna showed in a model of signal transduction that good predictions were possible despite only order of magnitude estimates of parameter values \citep{Brown:2003aa}.
Sethna and coworkers later showed that model performance is often controlled by only a few parameter combinations, 
a characteristic seemingly universal to multi-parameter models referred to as \textit{sloppiness} \citep{Machta:2013aa}.
We have also demonstrated \textit{sloppy} behavior in a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa,Lequieu:2011aa}.
Thus, given our previous experience with models containing hundreds of unknown parameters, we expect parameter estimation to be a manageable challenge. 
On the other hand, a critical challenge will be the estimation of allosteric connectivity at a genome scale. 
The regulation of glycolytic enzymes, such as phosphofructokinase I, has been studied for many years \citep{BERG_BIOCHEMISTRY,Peskov:2008aa}. 
The allosteric regulation of metabolic enzymes can also be established from organism specific databases, such as EcoCyc \citep{Keseler:2013aa}, 
or more general allosteric databases, such as the AlloSteric Database \citep{Huang:2014aa}. 
However, for those enzymes that have not been well studied, we will need to infer allosteric interactions from experimental data.
In general, the reverse engineering of regulatory network structure from data is a very difficult problem. 
Recently, Sauer and colleagues have developed a systematic, model-based approach for the identification of allosteric regulation \textit{in vivo} \cite{Link:2013aa}. 
They tested the effects of many putative allosteric protein-metabolite interactions on the performance of a kinetic model of glycolysis against dynamic metabolomic and fluxomic measurements. 
A method similar to this may be easily applied to cell-free systems in order to identify relevant \textit{in vitro} allosteric interactions. 
Because omics measurements of cell-free environments are easy to obtain, identification of large-scale allosteric control structures may be possible. 
Also, there are many different approaches from the reverse engineering of gene regulatory networks that perhaps could be adopted to this problem, however this remains an open question.
Lastly, the choice of max/min integration rules or the particular form of the transfer functions could be generalized to include other rule types and functions. 
Theoretically, an integration rule is a function whose domain is a set of transfer function inputs, and whose range is $v\in[0,1]$.
Thus, integration rules other than max/min could be used, such as the mean or the product, assuming the range of the transfer functions is always $f\in[0,1]$.
Alternative integration rules such as the mean might have different properties which could influence model identification or performance. 
For example, a mean integration rule would be differentiable, which allows derivative-based optimization approaches to be used. 
The particular form of the transfer function could also be explored. We choose a Hill-like function because of its 
prominence in the systems and synthetic biology community. However, the only mathematical requirement for a transfer function is that it map a non-negative continuous or categorical variable into the 
range $f\in[0,1]$. Thus, many types of transfer functions are possible.

@@

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure-3-EnzymeKinetics.pdf}
\caption{Kinetics of simple transformations in the presence of activation and inhibition. 
\textbf{A}:The conversion of substrate $S$ to product $P$ by enzyme $E$ was activated by $S$. 
For a fixed control gain parameter $\kappa_{control}$, the reaction rate approached a step for increasing cooperativity control parameter $\eta$. 
\textbf{B}:The conversion of substrate $S$ to product $P$ by enzyme $E$ with inhibitor $I$. 
For a fixed control gain parameter $\kappa_{control}$, the reaction rate approximated non-competitive inhibition for increasing cooperativity control parameter $\eta$. 
}\label{fig-kinetics-simulations}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth,height=0.6\textheight]{./figs/Figure-8-Flux.pdf}
\caption{Metabolic flux and control variables as a function of network type and particle index at t = 100 min.
The control variables governing $E_{1},E_{3}$ and $E_{6}$ activity and the scaled metabolic flux were calculated for the positively (top), 
uniformly (middle) and negatively (bottom) biased particle swarms (N = 100). 
Blue denotes a low value, while red denotes a high value for the respective quantity being plotted.
The particles from each swarm were sorted based upon simulation error (low to high error). 
\textbf{A}: Model performance for the positively biased particle swarm as a function of particle index.
\textbf{B}: Model performance for the uniformly biased particle swarm as a function of particle index.
\textbf{C}: Model performance for the negatively biased particle swarm as a function of particle index. 
Models with significant control mismatch showed distinct control and flux patterns versus those models with the correct or closely related control policies.
In particular, models with the correct control policy showed stronger inhibition of $E_{1}$ activity, leading to decreased flux from S$\rightarrow$P$_{1}$.
Conversely, models with significant mismatch had increased $E_{1}$ activity, leading to an altered flux distribution. 
This is especially apparent in the negatively biased particle swarm. 
}\label{fig-flux-pattern}
\end{figure}

>>>>

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large Dynamic Modeling of Cell-Free Biochemical Networks using Effective Kinetic Models}}
\vspace{0.05in}
{\par \centering \large{Joseph A. Wayman, Adithya Sagar and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering \large{School of Chemical and Biomolecular Engineering}}
{\par \centering \large{Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~Effective models of metabolism}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\textit{Processes}}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Associate Professor, School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853} 
{\par \centering Email: jdv27@cornell.edu} 
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Cell-free systems offer many advantages for the study, manipulation and modeling of metabolism compared to \textit{in vivo} processes.
Many of the challenges confronting genome-scale kinetic modeling can potentially be overcome in a cell-free system. 
For example, there is no complex transcriptional regulation to consider, transient metabolic measurements are easier to obtain, and we no longer have to consider cell growth. 
Thus, cell-free operation holds several significant advantages for model development, identification and validation. 
Theoretically, genome-scale cell-free kinetic models may be possible for industrially important organisms, such as \textit{E. coli}, if a simple, tractable framework for integrating allosteric regulation with enzyme kinetics can be formulated. Toward this unmet need, we present an effective biochemical network modeling framework for building dynamic cell-free metabolic models.
The key innovation of our approach is the integration of simple effective rules encoding complex allosteric regulation with traditional kinetic pathway modeling. 
We tested our approach by modeling the time evolution of several hypothetical cell-free metabolic networks. 
We found that simple effective rules, when integrated with traditional enzyme kinetic expressions, captured complex allosteric patterns such as ultrasensitivity or non-competitive inhibition in the absence of mechanistic information. Second, when integrated into network models, these rules captured classic regulatory patterns such as product-induced feedback inhibition. 
Lastly, we showed, at least for the network architectures considered here, that we could simultaneously estimate kinetic parameters and allosteric connectivity from synthetic data 
starting from an unbiased collection of possible allosteric structures using particle swarm optimization. However, when starting with an initial population that was heavily enriched with incorrect structures, our particle swarm approach could converge to an incorrect structure. While only an initial proof-of-concept, the framework presented here could be an important first step toward genome-scale cell-free kinetic modeling of the biosynthetic capacity of industrially important organisms.

@@

\section*{Introduction}
Mathematical modeling has long contributed to our understanding of metabolism. 
Decades before the genomics revolution, mechanistically, structured metabolic models arose from the desire to predict microbial phenotypes resulting from changes in intracellular or extracellular states \citep{1976_fredrickson_BiotechBioeng}. 
The single cell \textit{E. coli} models of Shuler and coworkers pioneered the construction of large-scale, dynamic metabolic models that incorporated multiple, regulated catabolic and anabolic pathways constrained by experimentally determined kinetic parameters \citep{1984_domach_shuler_BiotechBioeng_01}. 
Shuler and coworkers generated many single cell kinetic models, including single cell models of eukaryotes \citep{1989_steinmeyer_shuler_ChemEngSci,1992_wu_shuler_AnnNYAcadSci}, minimal cell architectures \citep{2004_castellanos_shuler_PNAS}, as well as DNA sequence based whole-cell models of \textit{E. coli} \citep{2008_atlas_shuler_IETSysBio}.
Conversely, highly abstracted kinetic frameworks, such as the cybernetic framework, represented a paradigm shift, viewing cells as growth-optimizing strategists \citep{1985_dhurjati_ramkrishna_tsao_BiotechBioeng}. 
Cybernetic models have been highly successful at predicting metabolic choice behavior, e.g., diauxie behavior \citep{1986_kompala_ramkrishna_tsao_BiotechBioeng}, steady-state multiplicity \citep{2012_kim_ramkrishna_BiotechProg}, as well as the cellular response to metabolic engineering modifications \citep{1999_varner_ramkrishna_MetaEng}. 
Unfortunately, traditional, fully structured cybernetic models also suffer from an identifiability challenge, as both the kinetic parameters and an abstracted model of cellular objectives must be estimated simultaneously. 
However, recent cybernetic formulations from Ramkrishna and colleagues have successfully treated this identifiability challenge through elementary mode reduction, though the techniques replace detailed biological mechanism with an optimization heuristic \cite{2009_song_ramkrishna_BiotechBioeng,Song:2011aa}.

@@

In this study, we present an effective biochemical network modeling framework for building dynamic cell-free metabolic models.
The key innovation of our approach is the seamless integration of simple effective rules encoding complex regulation with traditional kinetic pathway modeling. 
This integration allows the description of complex regulatory interactions, such as time-dependent allosteric regulation of enzyme activity, in the absence of specific mechanistic information. 
The regulatory rules are easy to understand, easy to formulate and do not rely on overarching theoretical abstractions or restrictive assumptions. 
We tested our approach by modeling the time evolution of several hypothetical cell-free metabolic networks. 
In particular, we tested whether our effective modeling approach could describe classically expected enzyme kinetic behavior, and second whether we could simultaneously estimate kinetic parameters and regulatory connectivity, in the absence of specific mechanistic knowledge, from synthetic experimental data. 
Toward these questions, we explored five hypothetical cell-free networks. 
Each network shared the same enzymatic connectivity, but had different allosteric regulatory connectivity. 
We found that simple effective rules, when integrated with traditional enzyme kinetic expressions, captured complex allosteric patterns such as ultrasensitivity or non-competitive inhibition in the absence of mechanistic information. Second, when integrated into network models, these rules captured classical regulatory patterns such as product-induced feedback inhibition. 
Lastly, we showed, at least for the network architectures considered here, that we could simultaneously estimate kinetic parameters and allosteric connectivity from synthetic data 
starting from an unbiased collection of possible allosteric structures using particle swarm optimization. However, when starting with an initial population that was heavily enriched with incorrect structures, our particle swarm approach could converge to an incorrect structure.
While only an initial proof-of-concept, the framework presented here could be an important first step toward genome-scale cell-free kinetic modeling of the biosynthetic capacity of industrially important organisms.

@@

We modified our particle swarm identification strategy to simultaneously search over both kinetic parameters and putative control structures. 
In addition to our initial networks, we constructed three additional presumptive network models, each with the same enzymatic connectivity but different allosteric regulation of the pathway enzymes (Fig. \ref{fig-alternative-networks}). 
We then initialized a population of particles, each with one of the five potential regulatory programs and randomized kinetic parameters. 
Thus, we generated an initial population of particles that had \textit{both} different kinetic parameters as well as different control structures. 
We biased the distribution of the particle population according to our \textit{a prior} belief of the correct regulatory program. 
To this end, we considered three different priors, a uniform distribution where each putative regulatory structure represented 20\% of the population and two mixed distributions that were either positively or negatively biased towards the correct structure (network A). 
In both the positively biased and uniform cases the PSO clearly differentiated between the true or closely related structures and those that were materially different (Fig. \ref{fig-control-search}). 
As expected, the positively biased population (40\% of the initial particle population seeded with network A) gave the best results, where the correct structure was preferentially identified (Fig. \ref{fig-control-search}A). 
On the other hand, when given a uniform distribution, the PSO approach identified a combination of network A and network C as the most likely control structures (Fig. \ref{fig-control-search}B). 
Network A and C differ by the regulatory connection between the end product P$_2$ and enzyme $E_{1}$; in network A, end product P$_{2}$ was assumed to inhibit $E_{1}$, while in network C, end product P$_{2}$ activated $E_{1}$.  
Lastly, when the initial population was heavily biased towards incorrect structures (initial population seeded with 90\% incorrect structures), the particle swarm \textit{misidentified} the correct allosteric structure (Fig. \ref{fig-control-search}C). Interestingly, while each particle swarm identified parameter sets that minimized the simulation error, the estimated parameter values were not necessarily similar to the true parameters. 
The angle between the estimated and true parameters was not consistently small across the swarms (identical parameters would give an angle of zero). 
This suggested that our particle swarm approach identified a \textit{sloppy} ensemble, i.e., parameter estimates that were individually incorrect but collectively exhibited the correct model behavior.

@@

\section*{Discussion}
In this study, we presented an effective kinetic modeling strategy to dynamically simulate cell-free biochemical networks. 
Our proposed strategy integrated traditional kinetic modeling with an effective rules based approach to dynamically describe metabolic regulation and control. 
We tested this approach by developing kinetic models of hypothetical cell-free metabolic networks. 
In particular, we tested whether our effective modeling approach could describe classically expected behavior, and second whether we could simultaneously estimate kinetic parameters and regulatory connectivity, in the absence of specific mechanistic knowledge, from synthetic experimental data. 
Toward these questions, we explored five hypothetical cell-free networks. 
In each network, a substrate S was converted to the end products P$_{1}$ and P$_{2}$ through a series of enzymatically catalyzed reactions, including a branch point at a hypothetical metabolite M$_{2}$. 
Each network also included the same cofactors and cofactor recycle architecture. 
However, while all five networks shared the same enzymatic connectivity, each had different allosteric regulatory connectivity. 
We found that simple effective rules, when integrated with traditional enzyme kinetic expressions, could capture complex allosteric patterns such as ultrasensitivity, or non-competitive inhibition in the absence of specific mechanistic information. Moreover, when integrated into network models, these rules captured classical regulatory patterns such as product-induced feedback inhibition. 
Lastly, we simultaneously estimated kinetic parameters and discriminated between competing regulatory structures, using synthetic data in combination with a modified particle swarm approach.
If we considered all putative regulatory architectures to be equally likely, we were able to estimate a \textit{sloppy} ensemble of models with the correct architecture and kinetic parameters.
Thus, we identified parameter values that were different from their true values, but nonetheless produced reasonable model performance (low error). This suggested that we captured important parameter combinations (stiff combinations), while simultaneously missing other parameter combinations (sloppy combinations). This was similar to the earlier study of Brown and Sethna \citep{Brown:2003aa}, 
which showed that reasonable model predictions were possible, despite sometimes only order of magnitude parameter estimates, if the stiff parameter combinations were well constrained.

@@

There are several critical questions that should be explored following this proof-of-concept study.
It is unclear how parameter identification will scale to genome-scale networks, and second it is unclear how we will identify allosteric connectivity at a genome-scale. 
The enzymatic connectivity for genome-scale cell-free networks can easily be established by stripping away the growth and cell wall machinery from whole cell genome reconstructions.
Then metabolic fluxes can be transformed into kinetic expressions using heuristics such multiple saturation kinetics, which are then modified by our rule-based control variables.  
This leaves a large number of unknown kinetic constants that must be estimated from time-resolved metabolite measurements. 
Ensemble modeling is a well-established approach for parameter identification in large-scale deterministic models. 
Liao and coworkers developed a method that generates an ensemble of kinetic models that all approach the same steady-state, one determined by fluxomics measurements \cite{Tran:2008aa}. 
The best subpopulation of candidate models were selected based on their agreement with further measurements of genetically perturbed systems. 
Our work relies on heuristic search optimization to identify kinetic models consistent with steady-state and dynamic time-series measurements of cellular species \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa,Lequieu:2011aa}. 
Instead of estimating a single yet highly uncertain parameter set, both approaches estimate an ensemble of parameter sets whose model behavior recapitulates experimental measurements. 
Here, we showed that particle swarm optimization quickly identified an ensemble of model parameters, at least for proof-of-concept metabolic networks using synthetic data. 
This suggested that we can expect reasonable model predictions, despite only partial parameter knowledge, as network size grows if we have properly designed experiments. 
Brown and Sethna showed in a model of signal transduction that good predictions were possible despite only order of magnitude estimates of parameter values \citep{Brown:2003aa}.
Sethna and coworkers later showed that model performance is often controlled by only a few parameter combinations, 
a characteristic seemingly universal to multi-parameter models referred to as \textit{sloppiness} \citep{Machta:2013aa}.
We have also demonstrated \textit{sloppy} behavior in a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa,Lequieu:2011aa}.
Thus, given our previous experience with models containing hundreds of unknown parameters, we expect parameter estimation to be a manageable challenge.

A critical challenge will be the estimation of allosteric connectivity at a genome scale. 
The regulation of glycolytic enzymes, such as phosphofructokinase I, has been studied for many years \citep{BERG_BIOCHEMISTRY,Peskov:2008aa}. 
The allosteric regulation of metabolic enzymes can also be established from organism specific databases, such as EcoCyc \citep{Keseler:2013aa}, 
or more general allosteric databases, such as the AlloSteric Database \citep{Huang:2014aa}. 
However, for those enzymes that have not been well studied, we will need to infer allosteric interactions from experimental data.
In general, the reverse engineering of regulatory network structure from data is a very difficult problem. 
Recently, Sauer and colleagues have developed a systematic, model-based approach for the identification of allosteric regulation \textit{in vivo} \cite{Link:2013aa}. 
They tested the effects of many putative allosteric protein-metabolite interactions on the performance of a kinetic model of glycolysis against dynamic metabolomic and fluxomic measurements. 
A method similar to this may be easily applied to cell-free systems in order to identify relevant \textit{in vitro} allosteric interactions. 
Because omics measurements of cell-free environments are easy to obtain, identification of large-scale allosteric control structures may be possible. 
Also, there are many different approaches from the reverse engineering of gene regulatory networks that perhaps could be adopted to this problem, however this remains an open question.
For example, one could imagine designing pulse chase experiments which maximally distinguish between competing allosteric models, similar to the earlier work of Kremling et al \citep{Kremling:2004aa}, or
iteratively estimate model structures similar to Doyle and coworkers \citep{Gadkar:2005aa}.
Lastly, the choice of max/min integration rules or the particular form of the transfer functions could be generalized to include other rule types and functions. 
Theoretically, an integration rule is a function whose domain is a set of transfer function inputs, and whose range is $v\in[0,1]$.
Thus, integration rules other than max/min could be used, such as the mean or the product, assuming the range of the transfer functions is always $f\in[0,1]$.
Alternative integration rules such as the mean might have different properties which could influence model identification or performance. 
For example, a mean integration rule would be differentiable, which allows derivative-based optimization approaches to be used. 
The particular form of the transfer function could also be explored. We choose a Hill-like function because of its 
prominence in the systems and synthetic biology community. However, the only mathematical requirement for a transfer function is that it map a non-negative continuous or categorical variable into the 
range $f\in[0,1]$. Thus, many types of transfer functions are possible.

@@

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure-3-EnzymeKinetics.pdf}
\caption{Kinetics of simple transformations in the presence of activation and inhibition. 
\textbf{A}:The conversion of substrate $S$ to product $P$ by enzyme $E$ was activated by $S$. 
For a fixed control gain parameter $\kappa_{control}$, the reaction rate approached a step for increasing cooperativity control parameter $\eta$. 
For activation simulations $\kappa_{control}=$0.05 and $\eta=\left\{0.01,0.1,1,2,4,6,8,10\right\}$.
\textbf{B}:The conversion of substrate $S$ to product $P$ by enzyme $E$ with inhibitor $I$. 
For a fixed control gain parameter $\kappa_{control}$, the reaction rate approximated non-competitive inhibition for increasing cooperativity control parameter $\eta$. 
For the inhibition simulations $\kappa_{control}=$1.5 and $\eta=\left\{0.01,0.1,1,2,4,6,8,10\right\}$.}\label{fig-kinetics-simulations}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth,height=0.6\textheight]{./figs/Figure-8-Flux.pdf}
\caption{Metabolic flux and control variables as a function of network type and particle index at t = 100 min.
The particle error, the control variables governing $E_{1},E_{3}$ and $E_{6}$ activity ($v_{1},v_{3}$ and $v_{3}$) and the scaled metabolic flux were calculated for the positively (top), 
uniformly (middle) and negatively (bottom) biased particle swarms (N = 100). 
Blue denotes a low value, while red denotes a high value for the respective quantity being plotted.
The particles from each swarm were sorted based upon simulation error (low to high error). 
\textbf{A}: Model performance for the positively biased particle swarm as a function of particle index.
\textbf{B}: Model performance for the uniformly biased particle swarm as a function of particle index.
\textbf{C}: Model performance for the negatively biased particle swarm as a function of particle index. 
Models with significant control mismatch showed distinct control and flux patterns versus those models with the correct or closely related control policies.
In particular, models with the correct control policy showed stronger inhibition of $E_{1}$ activity, leading to decreased flux from S$\rightarrow$P$_{1}$.
Conversely, models with significant mismatch had increased $E_{1}$ activity, leading to an altered flux distribution. 
This is especially apparent in the negatively biased particle swarm. 
}\label{fig-flux-pattern}
\end{figure}
